---
title: "Developing Data Products Shiny App Assignment"
author: "Yigit Ozan Berk"
date: "10/8/2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r initiation, message = FALSE, warning= FALSE, echo = FALSE}
library(ISLR)
```

## Introduction

- This app uses the Credit data set from ISLR package. 
- It analyzes the relationships between Income and other predictors. 
- It presents an input button for "student" variable indicator, and two sliders
for "balance" and "rating" variable inputs for prediction.
- It plots the residuals of the regression model of 
formula = Income ~ Rating*Balance + Student


## Exploratory Analysis

The model is selected from the analysis of predictors in the Credit dataset.
```{r pairs, echo = TRUE}
pairs(Credit)
```

## Selecting Predictors

The model begins with the most significant predictors, 'StudentYes', 'Balance' and 'Limit'. Upon further examination, because 'Rating' and 'Limit' is very highly correlated with a value of `r cor(Credit$Limit, Credit$Rating)` , 'Rating' predictor is used instead of 'Limit' because it makes more sense.
```{r pressure}
summary(lm(Income ~. -ID, data = Credit))
```


## Anova

Upon further examination of the variance, the 'Cards' predictor is not used because of unnecessary variance inflation, together with low F statistic.
```{r}
fit1 = lm(Income ~ Rating*Balance + Student, data = Credit)
fit2 = lm(Income ~ Rating*Balance + Student + Cards, data = Credit)
anova(fit1, fit2)
```



## Conclusion

- Using multivariate regression with the following formula 

```{r}
fit1
```
the user can get a good idea of the value of income of a person(student or non-student) given his rating and balance values.

- The app offers two sliders and an input button to prepare the model prediction value. 
- The app calculates and displays the prediction value after the values are submitted to the server by the submit button.