---
title: "my_fancy_presentation"
author: "Yigit Ozan Berk"
date: "10/3/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## My new slide

some content

***
A slide without a title

## My title

### A top level subheading
#### Second level subheading

## My title 2
- Bullet 1
- Bullet 2

1. item 1
2. item 2
3. plot of regression of mpg onto weight
```{r}
library(ggplot2)
qplot(x = wt, y = mpg, data = mtcars) + geom_smooth(method = "lm")
```

## my new slide 3

Plain text

*italicized text*

**Bold text**
`for (i %in% 1:10)`

```{r}
head(mtcars)
```

## my new slide 4

- to get rid of the # comment characters, use comment = ""
- you can use anything including nothing
- echo = TRUE shows the code as well

```{r, comment = "", echo = T}
head(mtcars)
```

## my new slide 5

- there are a lot of options in code chunks.


```{r, echo = F, fig.align='center', fig.cap="my Awesome figure"}
#echo = F to hide the code
plot(mtcars$wt, mtcars$mpg, xlab = "Weight", ylab = "MPG")
```

## my new slide 6

seankross.github.io/slides/Developing_Data_Products/R_Markdown/R_Markdown.html

if you go to this page, the slide will be hosted by github. 

